\% Artigos processados: 105
\begin{description}

\item[[TRH22]]
    \textsf{Sofia Teixeira, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"EWVM, a Web Virtual Machine to Support Code Generation in Compiler Courses"}, 11th Symposium on Languages, Applications and Technologies (SLATE 2022), ISBN: 978-3-95977-245-7, \url{https://drops.dagstuhl.de/opus/volltexte/2022/16753}, 2022.

\item[[CardRam22]]
    \textsf{Hugo André Coelho Cardoso, José Carlos Ramalho}, \emph{"Synthetic Data Generation From JSON Schemas"}, 11th Symposium on Languages, Applications and Technologies (SLATE 2022), ISBN: 978-3-95977-245-7, \url{https://drops.dagstuhl.de/opus/volltexte/2022/16751}, 2022.

\item[[CR22c]]
    \textsf{José Carlos Ramalho, Luís Filipe Cunha}, \emph{"Fine-Tuning BERT Models to Extract Named Entities from Archival Finding Aids"}, Workshops and Doctoral Consortium of the 26th International Conference on Theory and Practice of Digital Libraries, Pádua, Itália, \url{http://ceur-ws.org/Vol-3246/08_Paper1.pdf}, 2022.

\item[[CR22b]]
    \textsf{José Carlos Ramalho, Luís Filipe Cunha}, \emph{"NER in Archival Finding Aids: Next Level"}, WorldCist'22 - 10th World Conference on Information Systems and Technologies, Budva, Montenegro, \url{https://doi.org/10.1007/978-3-031-04819-7_33}, 2022.


\item[[MCR21b]]
    \textsf{Paulo Martins, Leandro Costa, José Carlos Ramalho}, \emph{"Knowledge Graph of press clippings referring social minorities"}, Linked Archives 2021: proceedings of Linked Archives International Workshop 2021, ISSN: 1613-0073, \url{http://ceur-ws.org/Vol-3019/LinkedArchives_2021_paper_16.pdf}, 2021.

\item[[CunRam21b]]
    \textsf{Luís Filipe Cunha, José Carlos Ramalho}, \emph{"Towards Entity Linking, NER in archival finding aids"}, Linked Archives 2021: proceedings of Linked Archives International Workshop 2021, ISSN: 1613-0073, \url{http://hdl.handle.net/1822/74782}, 2021.

\item[[PBR21]]
    \textsf{Ricardo Barros Pereira, Miguel A. Brito, José Carlos Ramalho}, \emph{"Development of self-diagnosis tests system using a DSL for creating new test suites for integration in a cyber-physical system"}, 10th Symposium on Languages, Applications and Technologies (SLATE 2021), ISBN: 978-3-95977-202-0, ISSN: 2190-6807, \url{http://hdl.handle.net/1822/73520}, 2021.

\item[[CR21]]
    \textsf{Filipa Alves dos Santos, Hugo André Coelho Cardoso, João da Cunha e Costa, Válter Ferreira Picas Carvalho, José Carlos Ramalho}, \emph{"DataGen: JSON/XML Dataset Generator"}, 10th Symposium on Languages, Applications and Technologies (SLATE 2021), ISBN: 978-3-95977-202-0, ISSN: 2190-6807, \url{http://hdl.handle.net/1822/73506}, 2021.

\item[[CunRam21]]
    \textsf{Luís Filipe Cunha, José Carlos Ramalho}, \emph{"NER in Archival Finding Aids"}, 10th Symposium on Languages, Applications and Technologies (SLATE 2021), ISBN: 978-3-95977-202-0, ISSN: 2190-6807, \url{http://hdl.handle.net/1822/73504}, 2021.

\item[[MCR21]]
    \textsf{José Carlos Ramalho, Paulo Martins, Leandro Costa}, \emph{"Major Minors - Ontological Representation of Minorities by Newspapers"}, 10th Symposium on Languages, Applications and Technologies (SLATE 2021), ISBN: 978-3-95977-202-0, ISSN: 2190-6807, \url{https://drops.dagstuhl.de/opus/volltexte/2021/14420}, 2021.

\item[[slate2020-OR]]
    \textsf{Francisco Oliveira, José Carlos Ramalho}, \emph{"Open web ontobud: An open source RDF4J frontend"}, SLATE 2020 - Symposium on Languages, Applications and Technologies, Escola Superior de Tecnologia Instituto Politécnico do Cávado e do Ave, Portugal, \url{http://hdl.handle.net/1822/71104}, 07.14, 2020.

\item[[slate2020-MR]]
    \textsf{Ezequiel Moreira, José Carlos Ramalho}, \emph{"SPARQLing Neo4J"}, SLATE 2020 - Symposium on Languages, Applications and Technologies, Escola Superior de Tecnologia Instituto Politécnico do Cávado e do Ave, Portugal, \url{http://hdl.handle.net/1822/71105}, 07.14, 2020.

\item[[cisti2017]]
    \textsf{José Carlos Ramalho, André Pereira, Miguel Ferreira, Luis Faria}, \emph{"RODA-in: A generic tool for the mass creation of Submission Information Packages"}, 12th Iberian Conference on Information Systems and Technologies (CISTI),, ISCTE, Lisboa, Portugal, \url{http://hdl.handle.net/1822/46093}, 06.24, 2017.

\item[[iPRES2016]]
    \textsf{Bruno Ferreira, Luis Faria, José Carlos Ramalho, Miguel Ferreira}, \emph{"Database Preservation Toolkit: A relational database conversion and normalization tool"}, iPRES: 13th International Conference on Digital Preservation, Bern, Switzerland, \url{http://hdl.handle.net/1822/43479}, 10.04, 2016.

\item[[bad2015b]]
    \textsf{André Santos, José Carlos Ramalho, Miguel Ferreira, Hélder Silva, Luis Faria, João Moreira, Eloy Rodrigues, José Carvalho}, \emph{"PT-CRIS: Um miradouro sobre o universo científico nacional"}, 12º Congresso Nacional BAD, Universidade de Évora, Portugal, \url{http://hdl.handle.net/1822/37856}, 10.21, 2015.

\item[[bad2015a]]
    \textsf{Alexandra Lourenço, José Carlos Ramalho, Pedro Penteado}, \emph{"Uma ontologia para os processos de negócio da Administração"}, 12º Congresso Nacional BAD, Universidade de Évora, Portugal, \url{http://hdl.handle.net/1822/38497}, 10.21, 2015.

\item[[edicic2015]]
    \textsf{Alexandra Lourenço, José Carlos Ramalho, Pedro Penteado}, \emph{"Da gestão da informação à gestão do conhecimento: uma proposta para a e-Administração em Portugal"}, VII Encuentro Ibérico EDICIC 2015, Madrid, Espanha, \url{http://hdl.handle.net/1822/38496}, 11.16, 2015.

\item[[dlm2014]]
    \textsf{José Carlos Ramalho, Luis Faria, Hélder Silva}, \emph{"Database Preservation Toolkit: a flexible tool to normalize and give access to databases"}, DLM Forum – 7th Triennial Conference: Making the Information Governance Landscape in Europe, Lisbon, Portugal, \url{http://hdl.handle.net/1822/35183}, 11.12, 2014.


\item[[ipres2013]]
    \textsf{Luis Faria, Alan Akbik, Barbara Sierman, Marcel Ras, Miguel Ferreira, José Carlos Ramalho}, \emph{"Automatic preservation watch using information extraction on the Web : a case study on semantic extraction of natural language for digital preservation"}, iPRES 2013 - 10th International Conference on Preservation of Digital Objects, Lisbon, Portugal, \url{http://hdl.handle.net/1822/25214}, 09.20, 2013.

\item[[or2013]]
    \textsf{Luis Faria, Christoph Becker, Kresimir Duretec , Miguel Ferreira, José Carlos Ramalho}, \emph{"Supporting the preservation lifecycle in repositories"}, Open Repositories 2013, Charlottetown, Canada, \url{ http://hdl.handle.net/1822/25101}, 07.08, 2013.

\item[[archiving2013]]
    \textsf{Miguel Ferreira, José Carlos Ramalho, Luis Faria}, \emph{"	Guidelines for legacy repository migration"}, IS&T Archiving 2013, Washington DC, Estados Unidos, \url{http://hdl.handle.net/1822/23978}, 04.02-05, 2013.

\item[[guimaramus2012]]
    \textsf{José Carlos Ramalho}, \emph{"	MuseCode: à procura da linearidade na música tonal"}, Pensar a Música, Palácio Vila Flor, Guimarães, \url{http://hdl.handle.net/1822/21660}, 03.23, 2012.

\item[[bad2012]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Luis Faria}, \emph{"Boas Práticas na Migração de Repositórios: lições aprendidas com o CALM e o ARQBASE"}, 11º Congresso Nacional de Bibliotecários, Arquivistas e Documentalistas, Fundação Calouste Gulbenkian, Lisboa, \url{http://hdl.handle.net/1822/21611}, 10.20, 2012.

\item[[icadl2012]]
    \textsf{Luis Faria, Petar Petrov, Kresimir Duretec , Christoph Becker, Miguel Ferreira, José Carlos Ramalho}, \emph{"Design and architecture of a novel preservation watch system"}, 14th International Conference on Asia-Pacific Digital Libraries, ICADL 2012, Taipei, Taiwan, \url{http://hdl.handle.net/1822/21447}, 11.12, 2012.

\item[[ipres2012]]
    \textsf{Christoph Becker, Kresimir Duretec , Petar Petrov, Luis Faria, Miguel Ferreira, José Carlos Ramalho}, \emph{"Preservation watch : what to monitor and how"}, iPRES 2012 - 9th International Conference on Preservation of Digital Objects, Toronto, Canada, \url{http://hdl.handle.net/1822/20666}, 11.01, 2012.

\item[[FR11e]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"New dimension in relational database preservation: raising the abstraction level"}, iPRES 2011 - 8th International Conference on Preservation of Digital Objects, Singapore, \url{http://hdl.handle.net/1822/15859}, 11.04, 2011.

\item[[FR11d]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Relational Databases Conceptual Preservation"}, VLDL2011 - Fourth Workshop on Very Large Digital Libraries (In conjunction with the International Conference on Theory and Practice of Digital Libraries (TPDL2011)), Berlin, Germany, \url{http://hdl.handle.net/1822/16019}, 09.29, 2011.

\item[[FR11c]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Using Ontologies to Abstract Relational Databases Conceptual Model"}, EPIA2011 - 15th Portuguese Conference on Artificial Intelligence, Lisboa, Portugal, \url{http://hdl.handle.net/1822/16113}, 10.10, 2011.

\item[[FR11b]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Preservation of Relational Databases"}, MSKE2011- International Conference on Managing Services in the Knowledge Economy, Universidade Lusíada, Vila Nova de Famalicão, Portugal, \url{http://hdl.handle.net/1822/13704}, 07.13, 2011.

\item[[FR11]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Using Ontologies in Database Preservation"}, 9th Conference on XML: Aplicações e Tecnologias Associadas, XATA 2011, ESEIG -IPP, Vila do Conde - Portugal, \url{http://hdl.handle.net/1822/13703}, 06.01, 2011.

\item[[FR10]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Significant Properties in the Preservation of Relational Databases"}, Research and Advanced Technology for Digital Libraries, 14th European Conference, ECDL2010, Glasgow, UK, ISBN: 978-3-642-15463-8, \url{http://hdl.handle.net/1822/13702}, 09.06, 2010.

\item[[RQR10]]
    \textsf{Maria Jesus Colmenero Ruiz, Miguel Ángel Marzal Garcia-Quismondo, José Carlos Ramalho}, \emph{"Criação e Aplicação de um modelo baseado em Topic Maps à Documentação Educativa"}, 10º Congresso Nacional de Bibliotecários, Arquivistas e Documentalistas, Guimarães, Portugal, ISBN: 978-972-9067-39-6, \url{http://hdl.handle.net/1822/11081}, 04.08, 2010.

\item[[FFR10]]
    \textsf{Miguel Ferreira, Luis Miguel Ferros, José Carlos Ramalho}, \emph{"DigitArq e o novo módulo de interoperabilidade OAI-PMH"}, 10º CONGRESSO NACIONAL DE BIBLIOTECÁRIOS, ARQUIVISTAS E DOCUMENTALISTAS, Guimarães, Portugal, \url{http://hdl.handle.net/1822/10532}, 04.08, 2010.

\item[[FR09]]
    \textsf{Ricardo Freitas, José Carlos Ramalho}, \emph{"Relational databases digital preservation"}, INFORUM : SIMPÓSIO DE INFORMÁTICA, Lisboa, 2009, Lisboa - Portugal, ISBN: 978-972-9348-18-1, \url{http://hdl.handle.net/1822/9740}, 11.09, 2009.

\item[[GARH09]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Topic maps constraint languages : understanding and comparing"}, International Journal of Reasoning-based Intelligent Systems, ISSN: 1755-0556. 1:3-4 (2009) 173 - 181, Journal: International Journal of Reasoning-based Intelligent Systems, Publisher: Inderscience, \url{http://hdl.handle.net/1822/9741}, 2009.

\item[[LRHGT08]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques, Jonas Gassen, Rogério Correa Turchetti}, \emph{"A Framework to specify, extract and manage Topic Maps driven by ontologies"}, SIGDOC'08 -- 26th ACM International Conference on Design of Communication, Lisboa, Portugal, \url{http://hdl.handle.net/1822/8236}, 09-23, 2008.

\item[[BKRFBR08]]
    \textsf{Christoph Becker, Michael Kraxner, Andreas Rauber, Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"Distributed Preservation Services: Integrating Planning and Actions"}, ECDL2008 - European Conference on Research and Advanced Technology for Digital Libraries, Aarhus, Danmark, \url{http://hdl.handle.net/1822/8239}, 09, 2008.

\item[[RFFCBC08]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Luis Faria, Rui Castro, Francisco Barbedo, Leandro Costa}, \emph{"RODA and Crib: A Service-Oriented Digital Repository"}, iPRES 2008: The Fifth International Conference on Preservation of Digital Objects, London, England, \url{http://hdl.handle.net/1822/8226}, 09, 2008.

\item[[RLA08]]
    \textsf{}, \emph{"XATA2008 (6ª Conferência Nacional) - XML: Aplicações e Tecnologias Associadas"}, Departamento de Informática da Universidade de Évora, Évora - Portugal, ISBN: 978-972-99166-5-6, \url{http://hdl.handle.net/1822/7708}, 02.14, 2008.

\item[[LMMVRH08-XATA]]
    \textsf{Giovani Rubert Librelotto, Mirkos Martins, Henrique Machado, Juliana Vizzotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Generating a Semantic Network for PubMed"}, XATA 2008 -  XML: Aplicações e Tecnologias Associadas (6ª Conferência Nacional), Évora - Portugal, 02.14, 2008.

\item[[FRF-XATA]]
    \textsf{Luis Miguel Ferros, José Carlos Ramalho, Miguel Ferreira}, \emph{"Creating a National Federation of Archives using OAI-PMH"}, XATA 2008 -  XML: Aplicações e Tecnologias Associadas (6ª Conferência Nacional), Évora - Portugal, \url{http://hdl.handle.net/1822/7821}, 02.14, 2008.

\item[[LGRH07-CLEI]]
    \textsf{Giovani Rubert Librelotto, Jonas Gassen, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Using Web Services to put Metamorphosis on the Web"}, XXXIII Conferencia Latinoamericana de Informática (CLEI2007), San José - Costa Rica, \url{http://hdl.handle.net/1822/7354}, 2007.

\item[[LATRH07-CLEI]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, Rogério Correa Turchetti, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Implementing TMCL: XTche - a topic map schema and constraint specification language"}, XXXIII Conferencia Latinoamericana de Informática (CLEI2007), San José - Costa Rica, \url{http://hdl.handle.net/1822/7364}, 2007.

\item[[LARH07-EPIA]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Topic Maps Constraint Languages: understanding and comparing"}, Portuguese Conference on Artificial Intelligence (EPIA2007), Guimarães - Portugal, \url{http://hdl.handle.net/1822/7368}, 2007.

\item[[LMMFRH07-TMRA]]
    \textsf{Giovani Rubert Librelotto, H. T. Machado, M. O. Martins, P. G. D. Ferreira, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Comparing Topic Maps Constraint Specification Languages"}, International Conference on Topic Maps Research and Applications (TMRA2007), Leipzig - Germany, \url{http://hdl.handle.net/1822/7399}, 2007.

\item[[LRH07-SEMINFO]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Estruturando o conhecimento com a norma ISO 13250 Topic Maps"}, VI Seminário de Informática - SEMINFO, 2007, Torres, RS, Brasil., Torres, RS - Brasil, 2007.

\item[[RFFC07-EML]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Luis Faria, Rui Castro}, \emph{"Relational database preservation through XML modelling"}, Extreme Markup Languages 2007, Montréal - Canada, August 2007., Montréal - Canada, \url{http://hdl.handle.net/1822/7120}, 08, 2007.

\item[[RLH07-EML]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Topic Maps applied to PubMed"}, Extreme Markup Languages 2007, Montréal - Canada, August 2007., Montréal - Canada, \url{http://hdl.handle.net/1822/7330}, 08, 2007.

\item[[LPRH07-XATA]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Navegando na Rede Semântica dos Topic Maps com o Ulisses"}, XATA 2007 -  XML: Aplicações e Tecnologias Associadas (5ª Conferência Nacional), Lisboa - Portugal, \url{http://hdl.handle.net/1822/6482}, 02.14, 2007.

\item[[LPRH07-TM]]
    \textsf{Giovani Rubert Librelotto, Renato Preigschadt de Azevedo, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Topic Maps Constraint Specification Languages: comparing AsTMa!, OSL, and XTche"}, Topic Maps 2007 -  The First International Topic Maps User Conference, Oslo - Norway, \url{http://hdl.handle.net/1822/6474}, 03.21, 2007.

\item[[FBR07-IJDL]]
    \textsf{Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"An intelligent decision support system for digital preservation"}, International Journal on Digital Libraries, ISSN: 1432-5012 (Print) 1432-1300 (Online), Journal: International Journal on Digital Libraries, Publisher: Springer Berlin / Heidelberg, \url{http://hdl.handle.net/1822/6648}, 05, 2007.

\item[[FCRBCF07]]
    \textsf{Miguel Ferreira, Rui Castro, José Carlos Ramalho, Francisco Barbedo, Luis Corujo, Luis Faria}, \emph{"RODA : Repositório de Objectos Digitais Autênticos"}, 9º CONGRESSO NACIONAL DE BIBLIOTECÁRIOS, ARQUIVISTAS E DOCUMENTALISTAS, Ponta Delgada, Açores, Portugal, \url{http://hdl.handle.net/1822/6330}, 03.29, 2007.

\item[[SFRL07]]
    \textsf{António Sousa, Luis Miguel Ferros, José Carlos Ramalho, Maria João Pires Lima}, \emph{"Consulta Real em Ambiente Virtual: implementação de uma sala de referência e leitura virtual num arquivo"}, 9º CONGRESSO NACIONAL DE BIBLIOTECÁRIOS, ARQUIVISTAS E DOCUMENTALISTAS, Ponta Delgada, Açores, Portugal, \url{http://hdl.handle.net/1822/6432}, 03.29, 2007.

\item[[RFCFBC07]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Rui Castro, Luis Faria, Francisco Barbedo, Luis Corujo}, \emph{"XML e preservação digital"}, XATA2007 : XML : aplicações e tecnologias associadas : actas da Conferência Nacional, Lisboa, Portugal, \url{http://hdl.handle.net/1822/6224}, 02.15, 2007.

\item[[RLC07]]
    \textsf{}, \emph{"XATA2007 (5ª Conferência Nacional) - XML: Aplicações e Tecnologias Associadas"}, Faculdade de Ciências - Universidade de Lisboa, Lisboa - Portugal, ISBN: 978-972-99166-4-9, \url{http://hdl.handle.net/1822/6234}, 02.14-15, 2007.

\item[[FBR07]]
    \textsf{Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"CRiB : preservation services for Digital Repositories"}, \url{http://hdl.handle.net/1822/6195}, \url{Apresentação efectuada na International Conference Open Repositories, 2, San Antonio, Texas, United States of America, 2007.}, 2007.

\item[[LRH06-SBSI06]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Metamorphosis: An Environment to Achieve Semantic Interoperability with Topic Maps"}, III Simpósio Brasileiro de Sistemas de Informação, Curitiba, Paraná - Brasil, \url{http://hdl.handle.net/1822/7359}, 2006.

\item[[RFCFBC06]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Rui Castro, Luis Faria, Francisco Barbedo, Luis Corujo}, \emph{"RODA - Desenvolvimento de uma plataforma tecnológica para suporte de Arquivos Digitais"}, \url{http://hdl.handle.net/1822/6420}, \url{Apresentação efectuada no Encontro Sofware Livre na Administração Pública, Torre do Tombo, Lisboa - Portugal, em 29 de Novembro de 2006.}, 2006.

\item[[RLS06]]
    \textsf{}, \emph{"XATA2006 (4ª Conferência Nacional) - XML: Aplicações e Tecnologias Associadas"}, Instituto Politécnico de Portalegre, Portalegre - Portugal, ISBN: 972-99166-2-4, \url{http://hdl.handle.net/1822/4744}, 02.09-10, 2006.

\item[[RFFLS06]]
    \textsf{José Carlos Ramalho, Miguel Ferreira, Luis Miguel Ferros, Maria João Pires Lima, António Sousa}, \emph{"Digitarq2 : nova arquitectura aplicacional para gestão de Arquivos Definitivos"}, 2ª Conferência Internacional de Arquivos Empresariais, Seixal, Portugal, \url{http://hdl.handle.net/1822/6038}, 11.27, 2006.

\item[[FBR06-Ariadne]]
    \textsf{Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"A foundation for automatic digital preservation"}, Ariadne, ISSN: 1361-3200, Journal: Ariadne, \url{http://hdl.handle.net/1822/5571}, July, 2006.

\item[[RLH06]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Lecture Notes in Computer Science"}, ISBN: 3-540-32527-1, Volume: 3873 / 2006, Publisher: Springer-Verlag GmbH, \url{10.1007/11676904_2}, 02, 2006.

\item[[FBR06-XATA]]
    \textsf{Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"CRIB : a service oriented architecture for digital preservation outsourcing"}, XATA - XML: Aplicações e Tecnologias Associadas, Portalegre - Portugal, ISBN: 972-99166-2-4, \url{http://hdl.handle.net/1822/4457}, 02.09, 2006.

\item[[RD06-XATA]]
    \textsf{José Carlos Ramalho, Luis Miguel Alves Domingues}, \emph{"Especificação e Geração Automática de Navegadores para Redes Semânticas Baseados em Interfaces Web"}, XATA - XML: Aplicações e Tecnologias Associadas, Portalegre - Portugal, ISBN: 972-99166-2-4, \url{http://hdl.handle.net/1822/4576}, 02.09, 2006.

\item[[LRH06-XATA]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Topic Maps Aplicados ao Sistema de Informação do Museu da Emigração"}, XATA - XML: Aplicações e Tecnologias Associadas, Portalegre - Portugal, ISBN: 972-99166-2-4, \url{http://hdl.handle.net/1822/4910}, 02.09, 2006.

\item[[RSL05]]
    \textsf{}, \emph{"XATA2005 (3ª Conferência Nacional) - XML: Aplicações e Tecnologias Associadas"}, Casa da Torre, Vila Verde - Portugal, ISBN: 972-99166-1-6, \url{http://hdl.handle.net/1822/865}, 02.10-11, 2005.

\item[[Ram05]]
    \textsf{José Carlos Ramalho}, \emph{"Document Composer: uma aplicação XML para extracção de informação de repositórios XML"}, XML: Aplicações e Tecnologias Associadas (XATA 2005), Braga, ISBN: 972-99166-1-6, \url{http://hdl.handle.net/1822/884}, 02.10, 2005.

\item[[LRH05]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Obtendo Interoperabilidade Semântica em Sistemas Heterogéneos de Informação com Metamorphosis"}, XML: Aplicações e Tecnologias Associadas (XATA 2005), Braga, ISBN: 972-99166-1-6, \url{http://hdl.handle.net/1822/1020}, 02.10, 2005.

\item[[LRH05b]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Constraining XML Topic Maps with XTche"}, XML: Aplicações e Tecnologias Associadas (XATA 2005), Braga, ISBN: 972-99166-1-6, \url{http://hdl.handle.net/1822/994}, 02.10, 2005.

\item[[RL05]]
    \textsf{José Carlos Ramalho, Carlos J. Feijó Lopes}, \emph{"Web Services: Aplicações Distribuídas sobre Protocolos Internet"}, FCA - Editora de Informática, ISBN: 972-722-421-0, Publisher: Colecção Web Pro, 03, 2005.

\item[[RLH05]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Constraining Topic Maps: A TMCL declarative implementation"}, Extreme Markup Languages 2005, Montréal - Canada, 1-5 August 2005., Montréal - Canada, \url{http://hdl.handle.net/1822/3117}, 08.04, 2005.

\item[[RLH05b]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Metamorphosis - a Topic Maps based Environment to Handle Heterogeneous Information Resources"}, TMRA'05 - International Workshop on  Topic Map Research and Applications , Leipzig - Germany, 10.06, 2005.

\item[[FBR05-CAPSI]]
    \textsf{Miguel Ferreira, Ana Alice Baptista, José Carlos Ramalho}, \emph{"Avaliação Automática de Migração em Redes Distribuídas de Conversores"}, 6ª Conferência da Associação Portuguesa de Sistemas de Informação, Bragança - Portugal, \url{http://hdl.handle.net/1822/3616}, 10.26, 2005.

\item[[RLH05-revAPSI17]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Geração automática de interfaces Web para Sistemas de Informação: Metamorphosis"}, Revista da Associação Portuguesa de Sistemas de Informação, ISSN: 0872-7031, Journal: Revista da Associação Portuguesa de Sistemas de Informação, Volume: 17, Publisher: APSI, \url{http://hdl.handle.net/1822/4365}, 2005.

\item[[LRH05-CLEI]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Ulisses: Um Navegador Conceptual para Topic Maps"}, XXXI Conferência Latinoamericana de Informática, Santiago de Cali - Colombia, ISBN: 958-670-422-X, \url{http://hdl.handle.net/1822/3532}, 10.10, 2005.

\item[[LR04]]
    \textsf{Carlos J. Feijó Lopes, José Carlos Ramalho}, \emph{"Web Services: Metodologias de Desenvolvimento"}, XML: Aplicações e Tecnologias Associadas (XATA 2004), Faculdade de Engenharia da Universidade do Porto - Porto, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/559}, 02.12, 2004.

\item[[FR04]]
    \textsf{Miguel Ferreira, José Carlos Ramalho}, \emph{"Aquisição e Armazenamento de Metainformação no Contexto de um Arquivo"}, XML: Aplicações e Tecnologias Associadas (XATA 2004), Faculdade de Engenharia da Universidade do Porto - Porto, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/614}, 02.12, 2004.

\item[[LRH04]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Um Extractor de Topic Maps a partir de Recursos Heterogéneos de Informação"}, XML: Aplicações e Tecnologias Associadas (XATA 2004), Faculdade de Engenharia da Universidade do Porto - Porto, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/613}, 02.12, 2004.

\item[[RTFR04]]
    \textsf{José Carlos Ramalho, Pedro Taveira, Ricardo Freitas, Vasco Rocha}, \emph{"Gerador de Web Services para cadeias de transformações de documentos XML"}, XML: Aplicações e Tecnologias Associadas (XATA 2004), Faculdade de Engenharia da Universidade do Porto - Porto, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/561}, 02.12, 2004.

\item[[MRH04]]
    \textsf{Ricardo Martins, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Comparação de Linguagens para descrição de Interfaces baseadas em XML"}, XML: Aplicações e Tecnologias Associadas (XATA 2004), Faculdade de Engenharia da Universidade do Porto - Porto, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/562}, 02.12, 2004.

\item[[RS04]]
    \textsf{}, \emph{"XATA2004 (1ª Conferência Nacional) - XML: Aplicações e Tecnologias Associadas"}, Faculdade de Engenharia do Porto - Portugal, ISBN: 972-99166-0-8, \url{http://hdl.handle.net/1822/602}, 02.12-13, 2004.

\item[[FR04b]]
    \textsf{José Carlos Ramalho, Miguel Ferreira}, \emph{"DIGITARQ - CREATING AND MANAGING A DIGITAL ARCHIVE"}, Electronic Publishing 2004 - Building Digital Bridges: linking cultures, commerce and science, Department of Information and Documentation Science at the University of Brasilia, in Brazil, ISBN: 85-88130-02-05, \url{http://hdl.handle.net/1822/615}, 06.23-26, 2004.

\item[[RLH04]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Adrian - E-Learning Content Production (creating online exams)"}, Electronic Publishing 2004 - Building Digital Bridges: linking cultures, commerce and science, Department of Information and Documentation Science at the University of Brasilia, in Brazil, ISBN: 85-88130-02-05, \url{http://hdl.handle.net/1822/608}, 06.23-26, 2004.

\item[[LSRH04]]
    \textsf{Giovani Rubert Librelotto, Weber Souza, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Using the Ontology Paradigm to integrate Information Systems"}, International Conference on Knowledge Engineering and Decision Support, Porto - Portugal, \url{http://hdl.handle.net/1822/584}, 07.21-23, 2004.

\item[[LRH04-Clei]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Extração de Topic Maps no Oveia: Especificação e Processamento"}, 30ma Conferencia Latinoamericana de Informática (CLEI2004), Sociedad Peruana de Computación, Peru, ISBN: 9972-9876-2-0, \url{http://hdl.handle.net/1822/594}, Sept., 2004.

\item[[RHL04-eLes]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques, Giovani Rubert Librelotto}, \emph{"ADRIAN: Sistema de Suporte à Produção de Conteúdos"}, Conferência eLes'04 - E-Learning no Ensino Superior, Universidade de Aveiro, Portugal, ISBN: 972-789-134-9, \url{http://hdl.handle.net/1822/733}, Outubro, 2004.

\item[[FR04-capsi]]
    \textsf{Miguel Ferreira, José Carlos Ramalho}, \emph{"DigitArq: Creating a Historical Digital Archive"}, 5ª Conferência da APSI - Associação Portuguesa dos Sistemas de Informação, Instituto Superior Técnico, Lisboa, Portugal, \url{http://hdl.handle.net/1822/718}, 11.03, 2004.

\item[[LRH04-XMLw]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XTche - A Language for Topic Maps Schema and Constraints"}, XML Conference and Exhibition 2004 Proceedings, Washington, USA, \url{http://hdl.handle.net/1822/681}, 12, 2004.

\item[[LRH04-WWS]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Achieving Semantic Interoperability in Heterogeneous Information Systems with Topic Maps"}, I  Workshop de Web Semântica (WWS2004), Brasília, Brasil, 10.22, 2004.

\item[[RS2003]]
    \textsf{José Carlos Ramalho, José Luis Santos}, \emph{"Electronic Publishing of ADB editions"}, Electronic Publishing 2003: From information to knowledge, Universidade do Minho - Guimarães - Portugal, \url{http://hdl.handle.net/1822/552}, 06.25, 2003.

\item[[RLH2003]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Ontology driven Websites - Metamorphosis: a framework to specify and manage ontology driven websites"}, Electronic Publishing 2003: From information to knowledge, Universidade do Minho - Guimarães - Portugal, \url{http://hdl.handle.net/1822/553}, 06.25, 2003.

\item[[RLH03]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Ontology driven websites with Topic Maps"}, III International Conference on Web Engineering (ICWE), Journal: III International Conference on Web Engineering (ICWE), Volume: Oviedo - Espanha, 07.15, 2003.

\item[[JLRH03]]
    \textsf{M. H.  Jacinto, Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XCSL: XML Constraint Specification Language"}, CLEI Electronic Journal,  ISSN 0717-5000. 2003. , Journal: CLEI Electronic Journal,  ISSN 0717-5000. 2003. , Volume: Best papers of XXVIII Conferencia Latinoamericana de Informática - CLEI, 2002, 2003.

\item[[LRH03]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XML Topic Map Builder: Specification and Generation"}, XATA 2003 - XML, Aplicações e Tecnologias Associadas, Braga - Portugal, \url{http://hdl.handle.net/1822/620}, 02.13, 2003.

\item[[Ram03]]
    \textsf{José Carlos Ramalho}, \emph{"O XML nos curriculae das licenciaturas em informática"}, XATA 2003 - XML, Aplicações e Tecnologias Associadas, Braga - Portugal, \url{http://hdl.handle.net/1822/554}, 02.13, 2003.

\item[[LRH03b]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"TM-Builder: Um Construtor de Ontologias baseado em Topic Maps"}, CLEI'03, La Paz, Bolívia, \url{http://hdl.handle.net/1822/555}, 09.29, 2003.

\item[[LRH03c]]
    \textsf{Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Geração automática de interfaces Web para sistemas de informação: Metamorphosis"}, COOPMedia 2003, ISEP, Porto, Portugal, \url{http://hdl.handle.net/1822/556}, 10.08, 2003.

\item[[SMR03]]
    \textsf{Daniel Edgar Pinto Soares, Maria da Conceição Vieira Mota, José Carlos Ramalho}, \emph{"Xexam: uma linguagem de suporte para exames online (e-learning)"}, XATA2003 - XML: Aplicações e Tecnologias Associadas, Vila Verde - Portugal, \url{http://hdl.handle.net/1822/557}, 02.13-14, 2003.

\item[[JRH03]]
    \textsf{M. H.  Jacinto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XML Templates for Constraints (XTC), Um Nível de Abstracção para Linguagens de Especificação de Restrições"}, XATA2003 - XML: Aplicações e Tecnologias Associadas, Vila Verde - Portugal, \url{http://hdl.handle.net/1822/4911}, 02.13-14, 2003.

\item[[RHLA03]]
    \textsf{}, \emph{"XATA2003 - XML: Aplicações e Tecnologias Associadas"}, Vila Verde - Portugal, \url{http://hdl.handle.net/1822/607}, 02.13-14, 2003.

\item[[RLH03b]]
    \textsf{José Carlos Ramalho, Giovani Rubert Librelotto, Pedro Rangel Henriques}, \emph{"Adrian: a platform for E-Learning Content Prodution"}, Second International Conference on Multimedia and ICTs in Education, Badajoz - Espanha, \url{http://hdl.handle.net/1822/558}, 12.04, 2003.

\item[[RH2002]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XML e XSL: da teoria à prática"}, FCA - Editora de Informática, Publisher: Colecção Tecnologias da Informação, 11, 2002.

\item[[JLRH02]]
    \textsf{M. H.  Jacinto, Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XCSL: XML Constraint Specification Language"}, CLEI'02, Montevideo, Uruguai,, \url{http://hdl.handle.net/1822/612}, 11, 2002.

\item[[JLRHb02]]
    \textsf{M. H.  Jacinto, Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Bidirectional Conversion between XML Documents and Relational Data Bases"}, 7th International Conference on CSCW in  Design, Rio de Janeiro - Brasil, \url{http://hdl.handle.net/1822/601}, 09.25, 2002.

\item[[JLRHc02]]
    \textsf{M. H.  Jacinto, Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Constraint Specification Languages: comparing XCSL, Schematron and XML-Schema"}, XML Europe'02, Barcelona - Espanha, \url{http://hdl.handle.net/1822/619}, 05.20, 2002.

\item[[JLRH02b]]
    \textsf{M. H.  Jacinto, Giovani Rubert Librelotto, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"XCSL tutorial"}, ACM Crossroads, Journal: ACM Crossroads, Volume: 8.5, Publisher: ACM Press, \url{http://hdl.handle.net/1822/610}, 2002.

\item[[Ram01]]
    \textsf{José Carlos Ramalho}, \emph{"Constraining Content: specification and processing"}, XML Europe'01, Berlin - Germany, \url{http://hdl.handle.net/1822/623}, 05, 2001.

\item[[Ram2000]]
    \textsf{José Carlos Ramalho}, \emph{"Anotação Estrutural de Documentos e sua Semântica"}, Universidade do Minho, \url{http://hdl.handle.net/1822/5625}, 07, 2000.

\item[[JHR00]]
    \textsf{M. H.  Jacinto, Pedro Rangel Henriques, José Carlos Ramalho}, \emph{"XML and Semantic Validation"}, Logic, Language and Information: Proceedings of the First Workshop on Logic and Language, Instituto de Lógica, Lenguaje e Información, Universidad de Sevilla, Espanha, \url{http://hdl.handle.net/1822/597}, 11.29, 2000.

\item[[RARH98]]
    \textsf{José Carlos Ramalho, José João Almeida, J. G. Rocha, Pedro Rangel Henriques}, \emph{"SGML Documents: Where does quality go?"}, Markup Languages: theory and practice , ISSN: 1099-6621, Journal: Markup Languages: theory and practice , Volume: journal, num. 1 vol. 1, MIT Press, \url{http://hdl.handle.net/1822/644}, 01, 1999.

\item[[Ram99]]
    \textsf{José Carlos Ramalho}, \emph{"Recovering Old Paper Historical Documents Through SGML/XML Modelling"}, Digital Resources for the Humanities - DRH99, London - England, 1999.

\item[[RH99]]
    \textsf{José Carlos Ramalho, José João Almeida}, \emph{"XML::DT - a Perl Down Translation Module"}, XML Europe'99, Granada - Espanha, \url{http://hdl.handle.net/1822/632}, 1999.

\item[[Ram99b]]
    \textsf{José Carlos Ramalho}, \emph{"EINE'99 - II Escola de Informática da Sociedade Brasileira de Computação"}, Publisher: Sociedade Brasileira de Computação (SBC), 11.23-26, 1999.

\item[[RAH1998]]
    \textsf{José Carlos Ramalho, José João Almeida, Pedro Rangel Henriques}, \emph{"Algebraic Specification of Documents"}, Theoretical Computer Science journal, volume 199, Journal: Theoretical Computer Science journal, volume 199, Volume: Elsevier, 06, 1998.

\item[[RH98]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Qualidade na Publicação Electrónica: como controlá-la?"}, Sistemas de Informação, Journal: Sistemas de Informação, Volume: journal, num. 9, pg 69-83, Associação Portuguesa de Sistemas de Informação, 1998.

\item[[RLH98]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques, A. R. Lopes}, \emph{"Generating SGML specific editors: from DTDs to Attribute Grammars"}, Markup Technologies'98, Chicago - USA, \url{http://hdl.handle.net/1822/637}, 12, 1998.

\item[[RHb98]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Beyond DTDs: Constraining Data Content"}, SGML Europe 98 Conference, Paris - France, \url{http://hdl.handle.net/1822/636}, 05, 1998.

\item[[RHHRAF98]]
    \textsf{J. G. Rocha, M. R. Henriques, Pedro Rangel Henriques, José Carlos Ramalho, José João Almeida, J. L. Faria}, \emph{"Adapting Museum Structures for the Web: No Changes Needed!"}, Museums and the Web: An International Conference, Toronto, Canada, \url{http://hdl.handle.net/1822/624}, 04.22, 1998.

\item[[RH98b]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Qualidade na Publicação Electrónica: como controlá-la?"}, Quatic 98: III Encontro Nacional para a Qualidade nas Tecnologias da Informação e Comunicação, Guimarães - Portugal, \url{http://hdl.handle.net/1822/621}, 1998.

\item[[RH98c]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Informatics and Archives: a present needed symbiosis"}, Workshop internacional Informação e documentos de arquivo em ambientes digitais, Porto - Portugal, 1998.

\item[[RH98d]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Os desafios colocados pela informação estruturada no contexto dos arquivos: o arquivo distrital de Braga"}, IX encontro nacional de Museus e Autarquias, Loures - Portugal , 1998.

\item[[Ram98]]
    \textsf{José Carlos Ramalho}, \emph{"Tutorial: Documentação Estruturada - Será que a minha empresa deve migrar para uma tecnologia baseada em SGML?"}, \url{tutorial at Terceiro Encontro Nacional para a Qualidade nas tecnologias de informação e comunicação, Guimarães - Portugal}, 1998.

\item[[LRH97]]
    \textsf{A. R. Lopes, José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"INES, an Environment for Structured Editions of SGML Documents"}, II Brasilian Symposium on Programming Languages (SBLP'97), Campinas-SP -- Brasil, \url{http://hdl.handle.net/1822/616}, 09, 1997.

\item[[RRAH97]]
    \textsf{José Carlos Ramalho, J. G. Rocha, José João Almeida, Pedro Rangel Henriques}, \emph{"SGML Documents: Where Does Quality Go?"}, SGML/XML'97, Washinton - USA, \url{http://hdl.handle.net/1822/617}, 12.08, 1997.

\item[[RAH96]]
    \textsf{José Carlos Ramalho, José João Almeida, Pedro Rangel Henriques}, \emph{"Document Semantics: two approaches"}, SGML96, Boston - USA, \url{http://hdl.handle.net/1822/622}, 12, 1996.

\item[[Ram96]]
    \textsf{José Carlos Ramalho}, \emph{"DSSSL: basic topics"}, Workshop: Document    Processing and SGML, Universidade do Minho, 09, 1996.

\item[[RH94]]
    \textsf{José Carlos Ramalho, Pedro Rangel Henriques}, \emph{"Uma experiência de utilização da Internet na gestão pedagógica e sua formalização"}, I Conferência Nacional WWW,    Informação Multimédia na Internet, Universidade do Minho, 07, 1995.

\item[[RAH95]]
    \textsf{José Carlos Ramalho, José João Almeida, Pedro Rangel Henriques}, \emph{"DAVID, Algebraic Specification of Documents"}, Algebraic Methods in Language Processing, Twente - Holanda, ISSN: 0929-0672, \url{http://hdl.handle.net/1822/618}, 12, 1995.

\item[[SRFP95]]
    \textsf{H. Santos, José Carlos Ramalho, J. M. Fernandes, A. J. Proença}, \emph{"A Heterogeneous Computer Vision Architecture: Implementation Issues"}, Computing Systems in Engineering, Journal: Computing Systems in Engineering, Volume: volume 6, número 4/5, páginas 401-408, 1995.

\item[[Ram93]]
    \textsf{José Carlos Ramalho}, \emph{"Um Compilador para o GLiTCH"}, 10, 1993.

\item[[VECPAR93]]
    \textsf{H. Santos, José Carlos Ramalho, J. M. Fernandes, A. J. Proença}, \emph{"A Heterogeneous Computer Vision Architecture: Implementation Issues"}, 1st International Meeting on Vector and Paralell Processing, 1993.
\end{description}

